---
title: "WGCNA Tutorial"
author: "Marian L Schmidt"
date: "`r format(Sys.time(), '%Y %B, %d')`"
output:
  html_document:
    code_folding: hide
    highlight: default
    keep_md: no
    theme: journal
    toc: yes
    toc_float:  
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```



# Data Input/Cleaning

[Data Input & Cleaning Tutorial Here](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-01-dataInput.pdf)

## Download Packages

```{r packages}
# Install Packages
#install.packages("HelpersMG")
#install.packages("WGCNA")

# The GO.db package will not let me load WGCNA, so let's download it
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("GO.db", version = "3.8")

# Load the packages
library(HelpersMG)
library(WGCNA)
library(tidyverse)

# Set the working directory
current_directory <- getwd()
paste("Working in directory:", current_directory)
```


## Load Data
```{r load-data}
# Load data
load("data/Chloroplasts_removed/ByLake_Filtering/5in10/muskegon/muskegon_5in10_physeqs.RData")
rare_muskegon_physeq_5in10_abs

# The following setting is important, do not omit.
options(stringsAsFactors = FALSE)

# Prepare CLR transformed OTU Table 
clr_otu <- read.csv("data/Chloroplasts_removed/ByLake_Filtering/5in10/muskegon/muskegon_clr_otu_5in10.csv") %>%
  tibble::column_to_rownames(var = "X") 
# Prepare Taxonomy Table 
clr_tax <- tax_table(rare_muskegon_physeq_5in10_abs)
# Prepare Metadata file 
clr_metadata <- 
  sample_data(rare_muskegon_physeq_5in10_abs) %>%
  mutate(norep_filter_name = paste(substr(Sample_16S, 1,4), substr(Sample_16S, 6, 9), sep=""),
         # duplicate Sample_16S column to move it to the rownames
         names = Sample_16S) %>%
  tibble::column_to_rownames(var = "names") 

# Create the phyloseq object
clr_physeq <- merge_phyloseq(otu_table(clr_otu, taxa_are_rows = FALSE), clr_tax, sample_data(clr_metadata))

# Read in the producutivity data
prod_data <- read.csv("data/production_data.csv") %>%
  dplyr::filter(fraction == "Free" & limnion == "Surface") %>%  # Select only rows that are free-living
  dplyr::select(names, tot_bacprod, SD_tot_bacprod) %>%         # Select relevant columns for total productivity
  mutate(tot_bacprod = round(tot_bacprod, digits = 2),          # Round to 2 decimals
         SD_tot_bacprod = round(SD_tot_bacprod, digits = 2) ) %>%      
  dplyr::rename(norep_filter_name = names) %>%                  # Rename to match other data frame
  arrange(norep_filter_name)

# Put together a dataframe with metadata and production data
meta_prod_dat <-
  prod_data %>%
  left_join(clr_metadata ,by = "norep_filter_name") %>%
  mutate(names = Sample_16S) %>%
  tibble::column_to_rownames(var = "names")

###### FINAL PHYLOSEQ OBJECT
prod_clr_physeq_orig <- subset_samples(clr_physeq, Sample_16S %in% meta_prod_dat$Sample_16S)
# Overwrite the sample data to include productivity data
prod_clr_physeq <- merge_phyloseq(otu_table(prod_clr_physeq_orig), tax_table(prod_clr_physeq_orig), sample_data(meta_prod_dat))
prod_clr_physeq

# Pull out the information and clean up the data
prod_clr_otu <- otu_table(prod_clr_physeq)
datOTUClr <- as.data.frame(prod_clr_otu)

# Metadata 
prod_clr_dat <- data.frame(sample_data(prod_clr_physeq))

# Select only the numeric traits
datContinuous <- 
  prod_clr_dat %>%
  dplyr::select(c(tot_bacprod, SD_tot_bacprod, Total.cells, Total.count.sd, 
                  HNA.cells, HNA.sd, LNA.cells, LNA.sd, Total_Sequences))
head(datContinuous)
```


## Data Cleaning 

Below is an analysis to find out whether there are outliers. Thankfully, there are none!
```{r clean-data, fig.height=9, fig.width=12}
# Check for OTUs and samples with too many missing values
good_otus <- goodSamplesGenes(datOTUClr, verbose = 3);
good_otus$allOK

# Search the data for outliers via clustering
#sampleTree <- hclust(dist(datOTUClr), method = "average")
#sampleTree

# Plot it 
#sizeGrWindow(12, 9) # Set plotting window to be 12 by 9 inches
#par(cex = 1, mar = c(0,4,2,0))
#plot(sampleTree, main = "Sample Clustering: No Outliers", 
#     sub = "", xlab = "", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)

# Clust contains the samples we want to keep
paste("There are", ntaxa(prod_clr_physeq), "genes in the dataset.")
paste("There are", nsamples(prod_clr_physeq), "samples in the dataset.")

# Re-cluster samples
sampleTree2 <- hclust(dist(datOTUClr), method = "average")

# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors <- numbers2colors(datContinuous, signed = FALSE);

# Plot the sample dendrogram and the colors underneath.
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(datContinuous),
                    main = "Sample dendrogram and trait heatmap")
```


## Save the data
```{r data-save}
save(prod_clr_physeq, datOTUClr, datContinuous, file = "data/WGCNA/WGCNA_data.RData")
```




# Automatic network construction & Model Detection

[Automatic Netrowk Construction Tutorial Here](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-02-networkConstr-auto.pdf)

## Choosing the soft-thresholding power: analysis of network topology
```{r soft-threshold, fig.width=9, fig.height=5}
# Choose a set of soft-thresholding powers
powers <- c(c(1:10), seq(from = 12, to=20, by=2))

# Call the network topology analysis function
sft <- pickSoftThreshold(datOTUClr, powerVector = powers, verbose = 5)

# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2))

# Set some parameters
cex1 = 0.9

# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], 
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n", 
     main = paste("Scale independence"))

text(sft$fitIndices[,1], -sign(sft$fitIndices[,3]) * sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")

# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5], 
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```


#### The soft threshold is 3!

## One-step network construction & module detection

```{r module-detection, fig.height=9, fig.width=12}
# WARNING:
    # When performing this function with your own data:
        # you will NEED TO CHANGE THE PARAMETER OPTIONS
        # Blocksize = maxBlockSize = tells the function how loarge the largest block can be that the reader's computer can handle. Default is 5,000 BUT this can break the later code. 

net <- blockwiseModules(datOTUClr, TOMType = "unsigned", verbose = 3,
                        reassignThreshold = 0, mergeCutHeight = 0.25,
                        numeicLabels = TRUE, pamRespectsDendro = FALSE, saveTOMs = TRUE,
                        
                        # CHANGE THESE PARAMETERS!
                        power = 3, deepSplit = 2, 
                        maxBlockSize = 400, minModuleSize = 10,
                        saveTOMFileBase = "data/WGCNA/prodOTU_TOM")

# Can pull out information from object
head(net$colors)
head(net$MEs)

#How many modules?
length(unique(net$colors))
table(net$colors)
sum(table(net$colors))
```

You can see that there are a total of 7 modules and that now each module is named a color. The numbers under the colors tells you how many OTUs were sorted into that module. Each OTU is in exactly 1 module, and you can see that if you add up all of the numbers from the various modules you get 482 (the number of OTUs that we limited our analysis to above).

```{r module-detection2, fig.height=9, fig.width=12}
# Save the data!
moduleLabels <- net$colors
moduleColors <- labels2colors(net$colors)
MEs <- net$MEs;
geneTree <- net$dendrograms[[1]];
save(MEs, moduleLabels, moduleColors, geneTree, file = "prodOTU-02-networkConstruction-auto.RData")

# open a graphics window
sizeGrWindow(12, 9)

# Convert labels to colors for plotting
mergedColors <- labels2colors(net$colors)
conTree <- net$dendrograms[[1]]

# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                    "Module colors",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Consensus gene dendrogram and module colors")

# ADVICE: See recutBlockwiseTrees function to modify the criteria without having to recomput the network and clustering dendogram.
```

**Figure:** Clustering dendrogram of genes, with dissimilarity based on topological overlap, together with assigned module colors.

# Relating modules to external clinical traits

[Relating Modules to Traits Tutorial Here](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-03-relateModsToExt.pdf)


## Quantifying module-trait associations
```{r module-trait-associations, fig.width=10, fig.height=6}
# define numbers of genes and samples
nOTUs <- ncol(datOTUClr)
nSamples <- nrow(datOTUClr)

# Recalculate MEs with color labels
MEs0 <- moduleEigengenes(datOTUClr, moduleColors)$eigengenes
MEs <- orderMEs(MEs0)
moduleTraitCor <- cor(MEs, datContinuous, use = "p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)

# PLOT
sizeGrWindow(10,6)
textMatrix <- paste(signif(moduleTraitCor, 2), "\n(", signif(moduleTraitPvalue, 1), ")", sep = "")
dim(textMatrix) <- dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3))

# Display the correlation values within a heatmap
labeledHeatmap(Matrix = moduleTraitCor, 
               xLabels = names(datContinuous),
               yLabels = names(MEs), 
               ySymbols = names(MEs), 
               colorLabels = FALSE, 
               colors = blueWhiteRed(50),
               textMatrix = textMatrix, 
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait Relationships"))
```

**Figure:** Relationships of consensus module eignegenes and traits of samples collected from Muskegon Lake. Each row in the table corresponds to a consensus module (of OTUs), and each column is a trait. Numbers in the table report the correlations of the corresponding module eigengenes and traits, with a p-value printed below in parentheses. The table is color coded by correlation according to the legend. 


We are interested in classifying the modules that are associated with bacterial production. So therefore, the trait of interest is _**tot_bacprod**_.

From the plot above, we can see that the following modules as significantly (pvalues in the parentheses):  

- Blue, correlation = -0.52  
- Black, correlation = -0.41





## Gene relationship to trait and important modules 

### Gene significance and module membership

**Goal:** Quantify associations of individual genes with trait of interest (tot_bacprod)

**How?**  
  1. Define OTU significance (**GS**, _originally called gene significance_) as the absolute value of the correlation between the gene and the trait.  
  2. For each module, also define a quantitative measure of module membership **(MM)** as the correlation of the module eigngene and the gene expression profile. *Allows us to quantify the similarity of all genes on the array to every module.*  
  

```{r module-membership}
# Define variable weight containing the weight column of datTrait
bacprod <- as.data.frame(datContinuous$tot_bacprod)
names(bacprod) <- "bacprod" # rename

# Names (colors of the modules
modNames <- substring(names(MEs), 3) # Remove the first two letters in string

# Calculate the correlations between modules
geneModuleMembership <- as.data.frame(WGCNA::cor(datOTUClr, MEs, use = "p"))

# What are the p-values for each correlation?
MMPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples))

# What's the correlation for the trait: bacterial production?
geneTraitSignificance <- as.data.frame(cor(datOTUClr, bacprod, use = "p"))

# What are the p-values for each correlation?
GSPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples = nSamples))

names(geneTraitSignificance) <- paste("GS.", names(bacprod), sep = "")
names(GSPvalue) <- paste("p.GS.", names(bacprod), sep = "")
```


## Intramodular analysis: ID genes with high GS and MM

Using GS and MM, let's identify genes with a high significance for bacterial production as well as high module membership in interesting modules. 


```{r GS-MM-corr, fig.height=12, fig.width=8}
# Fix the names so that they match the actual color
names(geneModuleMembership) <- substring(names(geneModuleMembership), 3) # Remove the first two letters in string

par(mfrow = c(4,2))  
  
# Initialize for loop
      # NEED: modNames
for (i in names(geneModuleMembership)) {
  
  # Pull out the module we're working on
  module <- i
  #print(module)   
  
  # Find the index in the column of the dataframe 
  column <- match(module, modNames)
  #print(column)
  
  # Pull out the Gene Significance vs module membership of black module
  moduleGenes = moduleColors == module
  #print(moduleGenes)
  
  # Make the plot
  verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]), 
                 abs(geneTraitSignificance[moduleGenes, 1]),
                 xlab = paste("Module Membership in", module, "module"),
                 ylab = "Gene significance for bacterial production",
                 main = paste("Module membership vs. gene significnace \n"),
                 cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)
}    

```

**Figure :** Above is a scatterplot Gene Significance (GS) for bacterial production vs. Module Membership (MM) in the each of the different modules. 

Modules with a high & significant correlations:  

  - Positive Correlations:  
    - blue: cor = 0.5, p = 4e-04  
    - brown: cor = 0.37, p = 0.029  
    - red: cor = 0.13, p = 0.039  
          
  - Negative correlation:  
    - yellow: cor = -0.4, p = 0.0025  


## Summary output of network analysis

From the plot above, it looks like the **blue module** has the largest positive correlation with bacterial production. 

Which bacterial taxa make up this module? 

We just identified modules with high association with the trait of interest (i.e. brown & red, but not magenta or midnightblue), now let's merge this information with gene annotation information and write out a file that summarizes the results. 

```{r alldata}
# Combine pval, taxonomy, module membership, and gene significance into one dataframe

# Prepare pvalue df
GSpval <- GSPvalue %>%
  tibble::rownames_to_column(var = "OTU")

# Prepare taxonomy df
tax_df <- tax_table(prod_clr_physeq) %>%
  data.frame() %>%
  tibble::rownames_to_column(var = "OTU") %>%
  rename(Domain = Rank1, Phylum = Rank2, Class = Rank3, Order = Rank4, Family = Rank5, Genus = Rank6, Species = Rank7)

# Prepare module membership df
# Modify column names
#names(geneModuleMembership) <- substring(names(geneModuleMembership), 3)

gMM_df <- geneModuleMembership %>%
  tibble::rownames_to_column(var = "OTU") %>%
  gather(key = "moduleColor", value = "moduleMemberCorr", -OTU) 

# Prepare gene significance df
GS_bacprod_df <- geneTraitSignificance %>%
  data.frame() %>%
  tibble::rownames_to_column(var = "OTU")

# Put everything together 
allData_df <- gMM_df %>%
  left_join(GS_bacprod_df, by = "OTU") %>%
  left_join(GSpval, by = "OTU") %>%
  left_join(tax_df, by = "OTU")

# Write a file 
write.csv(allData_df, file = "data/WGCNA/allData_df_WGCNA.csv")
```



# Network visualization using WGCNA functions

[Visualization Tutorial Here](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-05-Visualization.pdf)

## Visualizing gene netowrks

```{r see-gene-networks, fig.width=9, fig.height=9}
# Calculate topological overlap. 
# Calculate during module detection, but calculating again here:
dissTOM <- 1-TOMsimilarityFromExpr(datOTUClr, power = 3) 

# Transform with a power to make moderately strong connectsion more visible in the heatmap
plotTOM <- dissTOM^7

# Set diagnol to NA for a nicer plot
diag(plotTOM) <- NA

######################## ALL GENES - even though it looks like subsampling 
nSelect = 482
nGenes = ncol(datOTUClr)
nSamples = nrow(datOTUClr)

# For reproducibility, we set the random seed
set.seed(10);
select = sample(nGenes, size = nSelect);
selectTOM = dissTOM[select, select];
# There’s no simple way of restricting a clustering tree to a subset of genes, so we must re-cluster.
selectTree = hclust(as.dist(selectTOM), method = "average")
selectColors = moduleColors[select];
# Open a graphical window
sizeGrWindow(9,9)
# Taking the dissimilarity to a power, say 10, makes the plot more informative by effectively changing
# the color palette; setting the diagonal to NA also improves the clarity of the plot
plotDiss = selectTOM^7;
diag(plotDiss) = NA;
TOMplot(plotDiss, selectTree, selectColors, main = "Network heatmap plot, selected genes")
```



## Visualizing the network of eigengenes

```{r see-eigengenes}
# Reclaculate module eigengenes
MEs = moduleEigengenes(datOTUClr, moduleColors)$eigengenes

# Isolate weight from the clinical traits
bacprod <- as.data.frame(datContinuous$tot_bacprod)
names(bacprod) = "__bacprod"

# Add the weight to existing module eigengenes
MET <- orderMEs(cbind(MEs, bacprod))
names(MET) <- substring(names(MET), 3)
```


```{r dendogram, fig.width=6, fig.height=6}
# Plot the dendogram
sizeGrWindow(6,6)
par(cex = 1.0)
plotEigengeneNetworks(MET, "Eigengene dendogram", marDendro = c(0,4,2,0),
                      plotHeatmaps = FALSE)


# Plot the heatmap 
par(cex = 1.0)
plotEigengeneNetworks(MET, "Eigengene adjacency heatmap", marHeatmap = c(3,4,2,2),
                      plotDendograms = TRUE, xLabeles = 90)
```

**Figure:** Visualization of the eigengene network representing the relationships among the modules and the clinical trait weight. Left panel shows a hierarchical clustering dendrogram of the eigengenes in which the dissimilarity of eigengenes EI , EJ is given by 1 − cor(EI , EJ ). The heatmap in the right panel shows the eigengene adjacency AIJ = (1 + cor(EI , EJ ))/2.


# OTUs
```{r OTUs-module}
library(DT)

# OTUs from Final figure
net$colors[names(net$colors) == "Otu000173"]
net$colors[names(net$colors) == "Otu000025"]
net$colors[names(net$colors) == "Otu000369"]

# How many OTUs in each color?
table(net$colors)

# Pull out only the brown OTU
brown_OTUs <- names(net$colors[net$colors == "brown"])
length(brown_OTUs)
brownDF <- data.frame(brown_OTUs) %>%
  rename(OTU = brown_OTUs) %>%
  left_join(dplyr::filter(allData_df, moduleColor == "brown"), by = "OTU") 

# Show the data
datatable(brownDF, options = list(pageLength = 10))

# Pull out only the turquoise OTU
turquoise_OTUs <- names(net$colors[net$colors == "turquoise"])
length(turquoise_OTUs)

turqDF <- data.frame(turquoise_OTUs) %>%
  rename(OTU = turquoise_OTUs) %>%
  left_join(dplyr::filter(allData_df, moduleColor == "turquoise"), by = "OTU") 

# Show the data
datatable(turqDF, options = list(pageLength = 10))
```



***



