---
title: "June 16 Analysis"
author: "Marian L Schmidt"
date: "6/16/2017"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 2
---


```{r setup, include=FALSE}
# For width of code chunks and scroll bar 
options(width=250)

knitr::opts_chunk$set(eval = TRUE, 
                      echo = TRUE,
                      include = TRUE,
                      warning = FALSE,
                      collapse = FALSE,
                      message = FALSE,
                      error = TRUE,
                      fig.align = "center")
```



### Load the necessary libraries and set colors 
```{r load-libraries}
################################# LOAD LIBRARIES ############################################
library(tidyverse)
library(cowplot)
source("../OutliersRemoved/set_colors.R")
source("../OutliersRemoved/Functions.R")
```


### Load in the data
```{r load-data}
#################################### LOAD DATA ##############################################

# Read in the absolute abundance data 
absolute_otu <- read.table(file="../../data/Chloroplasts_removed/nochloro_absolute_otu.tsv", header = TRUE) # Absolute OTU abundance table 

# Read in the taxonomy data 
tax <- read.table(file="../../data/Chloroplasts_removed/nochloro_taxonomy_otu.tsv", header = TRUE) %>%
  tibble::rownames_to_column() %>%
  dplyr::rename(Kingdom = Rank1,
         Phylum = Rank2, 
         Class = Rank3,
         Order = Rank4,
         Family = Rank5,
         Genus = Rank6,
         Species = Rank7,
         OTU = rowname) # Fix the Taxonomy

# Replace the phylum Proteobacteria with the class level
Phylum <- as.character(tax$Phylum)
Class <- as.character(tax$Class)

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Proteobacteria"){
    Phylum[i] <- Class[i]
  } 
}

# Overwrite the Phylum level with the new phylum classification
tax$Phylum <- Phylum # Add the new phylum level data back to phy

# Read in the productivity and flow cytometry data 
productivity <- read.table(file = "../../data/Chloroplasts_removed/productivity_data.tsv", header = TRUE) # Metadata file
```

# HNA 

## HNA Pool: 0.1% Filtering

@prubbens performed an analysis where initial filtering was 0.1% (results in `Final/HNA_OutliersRemoved/hna_scores_abun0.001.csv`).  The pipeline is found in `Final/HNA_OutliersRemoved/analysis_final_HNA_LNA_outliersremoved.ipynb`.


```{r HNAotus, dependson = "load-data", fig.height=5, fig.width=10}
# Vector of 3 OTUs pulled out by Peter's model in the above file 
hna_otus <- read.csv("hna_scores_abun0.001.csv", header = FALSE) %>%   # Gather the 26 OTUs that peter pulled out in his analysis 
  dplyr::rename(OTU = V1, Corr = V2)                                                               # Rename columns to be more intuitive

# Make a vector of OTU names to pull them out of the larger dataset
hna_otu_names <- as.character(hna_otus$OTU)

# How many otus are there in the HNA pool?
length(hna_otu_names)

# What is the taxonomy of these 3 OTUs?
tax %>% dplyr::filter(OTU %in% hna_otu_names)

# Make it more reproducble below
pick_columns <- length(hna_otu_names) + 1

# Put all the data together into one dataframe with only the important OTUs
AbsAbund_hna_otus <-  absolute_otu %>%                                           # Start with the OTU table with absolute abundance counts 
  dplyr::select(one_of(hna_otu_names)) %>%                                       # Pull out only the relevant OTUs from the above OTU table
  tibble::rownames_to_column() %>%                                               # Change the sample names to be a column
  dplyr::rename(Sample_16S = rowname) %>%                                        # Rename the sample names column to match up with other data frames
  dplyr::left_join(productivity, by = "Sample_16S") %>%                          # Join 26 OTU absolute abundance counts with rest of metadata 
  dplyr::filter(Lake == "Muskegon" & Depth == "Surface") %>%                     # Filter out samples that have productivity data (Muskegon & Surface)
  dplyr::select(-c(Platform, samples, Lake)) %>%                                 # Remove unnecessary columns 
  mutate(Site = factor(Site, levels = c("MOT", "MDP", "MBR", "MIN"))) %>%        # Fix order of factor for lake station
  gather("OTU", "Abs_Abund", 2:pick_columns) %>%                                 # Gather only columns which represent OTU abs abundance counts, and put it in *long* format
  dplyr::left_join(tax, by = "OTU") %>%                                          # Add the taxonomic information for each OTU
  dplyr::filter(tot_bacprod < 90) %>%                                            # Remove outliers to match peter's analysis
  mutate(OTU_fraction_HNA = Abs_Abund/HNA.cells)                                 # Calculate the fraction that each individual OTU takes up within the HNA pool for each sample



# Plot the absolute abundance data for each OTU
ggplot(AbsAbund_hna_otus, aes(x = reorder(OTU, Phylum), y = log10(Abs_Abund), fill = Phylum, color = Phylum)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 3) +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  ylab("log10(Abundance)") +
  #facet_grid(.~Phylum, scale = "free") +
  ggtitle("HNA OTUs from HNA Prediction") + 
  guides(fill = guide_legend(ncol=4),
         color = guide_legend(ncol=4)) +
  theme(legend.position = "bottom", #c(0.33, 0.85),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))

# Plot the fraction of the HNA pool for each OTU
ggplot(AbsAbund_hna_otus, aes(x = reorder(OTU, Phylum), y = OTU_fraction_HNA, fill = Phylum, color = Phylum)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 3) +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  ylab("Fraction of HNA Cells") +
  ggtitle("HNA OTUs from HNA Prediction") + 
  guides(fill = guide_legend(ncol=3),
         color = guide_legend(ncol=3)) +
  theme(legend.position = c(0.7, 0.85),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))
```

## Proportion of HNA Pool?

#### What proportion of the HNA pool is taken up by the `r length(hna_otu_names)` OTUs?

```{r HNA-sum-abundance, fig.height=4, fig.width=3, fig.align='center'}
# Calculate the sum of the HNA pool, the max of the HNA pool, and the median/mean
frac_HNA_stats_AbsAbund_hna_otus <- AbsAbund_hna_otus %>%                        # Take the dataframe from above
  dplyr::select(Sample_16S, OTU, OTU_fraction_HNA, Abs_Abund, HNA.cells) %>%   # Select only relevant columns
  group_by(Sample_16S) %>%                                                     # Make a data frame for each of the individual samples
  summarise(sum_fracHNA = sum(OTU_fraction_HNA),                               # Calculate the sum of the fraction of each OTU from the HNA pool (total HNA pool represented by each OTU)
            max_fracHNA = max(OTU_fraction_HNA),                               # Calculate the max the the line above
            median_fracHNA = median(OTU_fraction_HNA),                         # Calculate the median of 2 lines above
            mean_fracHNA = mean(OTU_fraction_HNA),                             # Calculate the mean of 3 lines above
            sum_abs_abund = sum(Abs_Abund)) %>%                                # What's the sum of the absolute abundance of the 26 OTUs within each sample? 
  mutate(All_Samples = "AllSamps_26_OTUs")


# Plot the variation in the sum of the HNA fraction with points/boxplot
ggplot(frac_HNA_stats_AbsAbund_hna_otus, 
       aes(y = sum_fracHNA, x = All_Samples, color = "All_Samples", fill = "All_Samples")) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) + 
  geom_point(size = 3, position = position_jitterdodge()) +
  ggtitle("HNA OTUs from \nHNA Prediction") + 
  xlab("All Samples") + scale_color_manual(values = "black") +
  scale_fill_manual(values = "black") +
  geom_abline(intercept = 1, slope = 0, color = "grey") +                                   # Draw a line at 1 
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, max(frac_HNA_stats_AbsAbund_hna_otus$sum_fracHNA) + 0.2), 
                     breaks = seq(0,  max(frac_HNA_stats_AbsAbund_hna_otus$sum_fracHNA) + 0.2, by = 0.2)) +
  ylab("\n Sum(Abundance/HNA.cells)") + xlab("Sample") +
  theme(legend.position = "none", axis.text.x = element_blank())
```


## Sum OTUs vs HNA
```{r pred-HNA, fig.height=4, fig.width=4, fig.align='center'}
all_data <- inner_join(frac_HNA_stats_AbsAbund_hna_otus, productivity, by = "Sample_16S") %>%    # Combine calculated stats above with rest of metadata
  dplyr::select(-c(Platform, All_Samples, samples, Lake, Fraction, Depth)) %>%                  # Remove futile columns
  mutate(pred_totHNA_counts = sum_fracHNA*HNA.cells)                                            # Sanity Check:  Back calculate the number of HNA cells 

# Plot the correlation between the HNA counts from the flow cytometer and the sum of the counts from the HNA OTUs
ggplot(all_data, aes(x = sum_abs_abund, y= HNA.cells)) +
  geom_point(size = 3) + ylab("HNA Cell Count") + 
  xlab("Sum(Abs_Abund of HNA OTUs)") + 
  geom_abline(intercept = 0, slope = 1) +                                   # Draw a 1:1 line 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))
```

# LNA

# LNA Pool: 0.1% Filtering

@prubbens performed an analysis where initial filtering was 0.1% (results in `Final/HNA_OutliersRemoved/lna_scores_abun0.001.csv`).  The pipeline is found in `Final/HNA_OutliersRemoved/analysis_final_HNA_LNA_outliersremoved.ipynb`.


```{r LNAotus, dependson = "load-data", fig.height=6, fig.width=16}
# Vector of LNA OTUs 
lna_otus <- read.csv("lna_scores_abun0.001.csv", header = FALSE) %>%   # Gather the LNA OTUs that peter pulled out in his analysis 
  dplyr::rename(OTU = V1, Corr = V2)                                                               # Rename columns to be more intuitive

# Make a vector of OTU names to pull them out of the larger dataset
lna_otu_names <- as.character(lna_otus$OTU)

# How many otus are there in the LNA pool?
length(lna_otu_names)

# What is the taxonomy of these LNA OTUs?
tax %>% dplyr::filter(OTU %in% lna_otu_names)

# Make it more reproducble below
lna_pick_columns <- length(lna_otu_names) + 1

# Put all the data together into one dataframe with only the important OTUs
AbsAbund_lna_otus <-  absolute_otu %>%                                           # Start with the OTU table with absolute abundance counts 
  dplyr::select(one_of(lna_otu_names)) %>%                                       # Pull out only the relevant OTUs from the above OTU table
  tibble::rownames_to_column() %>%                                               # Change the sample names to be a column
  dplyr::rename(Sample_16S = rowname) %>%                                        # Rename the sample names column to match up with other data frames
  dplyr::left_join(productivity, by = "Sample_16S") %>%                          # Join LNA OTU absolute abundance counts with rest of metadata 
  dplyr::filter(Lake == "Muskegon" & Depth == "Surface") %>%                     # Filter out samples that have productivity data (Muskegon & Surface)
  dplyr::select(-c(Platform, samples, Lake)) %>%                                 # Remove unnecessary columns 
  mutate(Site = factor(Site, levels = c("MOT", "MDP", "MBR", "MIN"))) %>%        # Fix order of factor for lake station
  gather("OTU", "Abs_Abund", 2:lna_pick_columns) %>%                             # Gather only columns which represent OTU abs abundance counts, and put it in *long* format
  dplyr::left_join(tax, by = "OTU") %>%                                          # Add the taxonomic information for each OTU
  dplyr::filter(tot_bacprod < 90) %>%                                            # Remove outliers to match peter's analysis
  mutate(OTU_fraction_LNA = Abs_Abund/LNA.cells)                                 # Calculate the fraction that each individual OTU takes up within the HNA pool for each sample



# Plot the absolute abundance data for each OTU
ggplot(AbsAbund_lna_otus, aes(x = reorder(OTU, Phylum), y = log10(Abs_Abund), fill = Phylum, color = Phylum)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 3) +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  ylab("log10(Abundance)") +
 # facet_grid(.~Phylum, scale = "free") +
  ggtitle("LNA OTUs from LNA Prediction") + 
  guides(fill = guide_legend(ncol=5),
         color = guide_legend(ncol=5)) +
  theme(legend.position = "bottom", #c(0.33, 0.85),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))

# Plot the fraction of the LNA pool for each OTU
ggplot(AbsAbund_lna_otus, aes(x = reorder(OTU, Phylum), y = OTU_fraction_LNA, fill = Phylum, color = Phylum)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 3) +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  ylab("Fraction of LNA Cells") +
  ggtitle("LNA OTUs from LNA Prediction") + 
  guides(fill = guide_legend(ncol=3),
         color = guide_legend(ncol=3)) +
  theme(legend.position = c(0.8, 0.85),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))
```



## Proportion of LNA Pool?

#### What proportion of the LNA pool is taken up by the `{r length(lna_otu_names)}` LNA OTUs?

```{r lna-sum-abundance, fig.height=4, fig.width=3, fig.align='center'}
# Calculate the sum of the HNA pool, the max of the HNA pool, and the median/mean
frac_LNA_stats_AbsAbund_lna_otus <- AbsAbund_lna_otus %>%                        # Take the dataframe from above
  dplyr::select(Sample_16S, OTU, OTU_fraction_LNA, Abs_Abund, LNA.cells) %>%   # Select only relevant columns
  group_by(Sample_16S) %>%                                                     # Make a data frame for each of the individual samples
  summarise(sum_fracLNA = sum(OTU_fraction_LNA),                               # Calculate the sum of the fraction of each OTU from the HNA pool (total HNA pool represented by each OTU)
            max_fracLNA = max(OTU_fraction_LNA),                               # Calculate the max the the line above
            median_fracLNA = median(OTU_fraction_LNA),                         # Calculate the median of 2 lines above
            mean_fracLNA = mean(OTU_fraction_LNA),                             # Calculate the mean of 3 lines above
            sum_abs_abund = sum(Abs_Abund)) %>%                                # What's the sum of the absolute abundance of the 26 OTUs within each sample? 
  mutate(All_Samples = "AllSamps_OTUs")


# Plot the variation in the sum of the HNA fraction with points/boxplot
ggplot(frac_LNA_stats_AbsAbund_lna_otus, 
       aes(y = sum_fracLNA, x = All_Samples, color = "All_Samples", fill = "All_Samples")) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) + 
  geom_point(size = 3, position = position_jitterdodge()) +
  ggtitle("LNA OTUs from \n LNA Prediction") + 
  xlab("All Samples") + scale_color_manual(values = "black") +
  scale_fill_manual(values = "black") +
  geom_abline(intercept = 1, slope = 0, color = "grey") +                                   # Draw a line at 1 
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, 1.8), 
                     breaks = seq(0,  1.8 + 0.2, by = 0.2)) +
  ylab("\n Sum(Abundance/LNA.cells)") + xlab("Sample") +
  theme(legend.position = "none", axis.text.x = element_blank())
```


## Sum OTUs vs LNA
```{r pred-LNA, fig.height=4, fig.width=4, fig.align='center'}
all_lna_data <- inner_join(frac_HNA_stats_AbsAbund_lna_otus, productivity, by = "Sample_16S") %>%    # Combine calculated stats above with rest of metadata
  dplyr::select(-c(Platform, All_Samples, samples, Lake, Fraction, Depth)) %>%                  # Remove futile columns
  mutate(pred_totLNA_counts = sum_fracLNA*LNA.cells)                                            # Sanity Check:  Back calculate the number of HNA cells 

# Plot the correlation between the HNA counts from the flow cytometer and the sum of the counts from the LNA OTUs
ggplot(all_lna_data, aes(x = sum_abs_abund, y= LNA.cells)) +
  geom_point(size = 3) + ylab("LNA Cell Count") + 
  xlab("Sum(Abs_Abund of LNA OTUs)") + 
  geom_abline(intercept = 0, slope = 1) +                                   # Draw a 1:1 line 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))
```

Quite a nice correlation with the LNA OTUS!



# Overlap in OTUs

The number of OTUs that are:  

- Unique to the HNA pool: `r length(hna_otu_names)`  
- Unique to the LNA pool: `r length(lna_otu_names)`  
- Shared between both pools: `r length(intersect(hna_otu_names, lna_otu_names))`


```{r}
OTUs_in_both <- intersect(hna_otu_names, lna_otu_names); sort(OTUs_in_both)
```
 
 
# Correlation of OTUs with Productivity  


### No transformation of the absolute abundances

```{r otu-prod, fig.height=15, fig.width=18, eval = TRUE, echo = TRUE}
summary(lm(tot_bacprod ~ Abs_Abund/OTU , data = dplyr::filter(AbsAbund_hna_otus, OTU %in% OTUs_in_both)))

ggplot(dplyr::filter(AbsAbund_hna_otus, OTU %in% OTUs_in_both), 
       aes(y = tot_bacprod,x = Abs_Abund, fill = Phylum, color = Phylum)) +
  geom_point(size = 3) + 
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  xlab("Absolute Abundance (cells/mL)") +
  facet_wrap(~OTU, scale = "free_x") + 
  geom_smooth(method = "lm") +
  ylab("Total Production (ug C/L/hr)") + 
  theme(legend.position = "bottom",
        legend.title = element_blank())  
```


### log10 transform the abundances 
```{r log10-otu-prod, fig.height=15, fig.width=18, eval = TRUE, echo = TRUE}
# Remove OTUs samples that have an abundance of 0
summary(lm(tot_bacprod ~ log10(Abs_Abund)/OTU , data = dplyr::filter(AbsAbund_hna_otus, OTU %in% OTUs_in_both & Abs_Abund > 0)))

ggplot(dplyr::filter(AbsAbund_hna_otus, OTU %in% OTUs_in_both & Abs_Abund > 0), 
       aes(y = tot_bacprod,x = log10(Abs_Abund), fill = Phylum, color = Phylum)) +
  geom_point(size = 3) + 
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  xlab("log10(Abundance)") +
  facet_wrap(~OTU, scale = "free_x") + 
  geom_smooth(method = "lm") +
  ylab("Total Production (ug C/L/hr)") + 
  theme(legend.position = "bottom",
        legend.title = element_blank())  
```



