---
title: "June 16 Analysis"
author: "Marian L Schmidt"
date: "6/16/2017"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 2
---


```{r setup, include=FALSE}
# For width of code chunks and scroll bar 
options(width=250)

knitr::opts_chunk$set(eval = TRUE, 
                      echo = TRUE,
                      include = TRUE,
                      warning = FALSE,
                      collapse = FALSE,
                      message = FALSE,
                      error = TRUE,
                      fig.align = "center")
```



### Load the necessary libraries and set colors 
```{r load-libraries}
################################# LOAD LIBRARIES ############################################
library(tidyverse)
library(cowplot)
source("../OutliersRemoved/set_colors.R")
source("../OutliersRemoved/Functions.R")
```


### Load in the data
```{r load-data}
#################################### LOAD DATA ##############################################

# Read in the absolute abundance data 
absolute_otu <- read.table(file="../../data/Chloroplasts_removed/nochloro_absolute_otu.tsv", header = TRUE) # Absolute OTU abundance table 

# Read in the taxonomy data 
tax <- read.table(file="../../data/Chloroplasts_removed/nochloro_taxonomy_otu.tsv", header = TRUE) %>%
  tibble::rownames_to_column() %>%
  dplyr::rename(Kingdom = Rank1,
         Phylum = Rank2, 
         Class = Rank3,
         Order = Rank4,
         Family = Rank5,
         Genus = Rank6,
         Species = Rank7,
         OTU = rowname) # Fix the Taxonomy

# Replace the phylum Proteobacteria with the class level
Phylum <- as.character(tax$Phylum)
Class <- as.character(tax$Class)

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Proteobacteria"){
    Phylum[i] <- Class[i]
  } 
}

# Overwrite the Phylum level with the new phylum classification
tax$Phylum <- Phylum # Add the new phylum level data back to phy

# Read in the productivity and flow cytometry data 
productivity <- read.table(file = "../../data/Chloroplasts_removed/productivity_data.tsv", header = TRUE) # Metadata file
```


# HNA Pool: 0.1% Filtering

@prubbens performed an analysis where initial filtering was 0.1% (results in `Final/HNA_OutliersRemoved/hna_scores_abun0.001.csv`).  The pipeline is found in `Final/HNA_OutliersRemoved/analysis_final_HNA_LNA_outliersremoved.ipynb`.


```{r HNAotus, dependson = "load-data", fig.height=4, fig.width=10}
# Vector of 3 OTUs pulled out by Peter's model in the above file 
hna_otus <- read.csv("hna_scores_abun0.001.csv", header = FALSE) %>%   # Gather the 26 OTUs that peter pulled out in his analysis 
  dplyr::rename(OTU = V1, Corr = V2)                                                               # Rename columns to be more intuitive

# Make a vector of OTU names to pull them out of the larger dataset
hna_otu_names <- as.character(hna_otus$OTU)

# How many otus are there in the HNA pool?
length(hna_otu_names)

# What is the taxonomy of these 3 OTUs?
tax %>% dplyr::filter(OTU %in% hna_otu_names)

# Make it more reproducble below
pick_columns <- length(hna_otu_names) + 1

# Put all the data together into one dataframe with only the important OTUs
AbsAbund_hna_otus <-  absolute_otu %>%                                           # Start with the OTU table with absolute abundance counts 
  dplyr::select(one_of(hna_otu_names)) %>%                                       # Pull out only the relevant OTUs from the above OTU table
  tibble::rownames_to_column() %>%                                               # Change the sample names to be a column
  dplyr::rename(Sample_16S = rowname) %>%                                        # Rename the sample names column to match up with other data frames
  dplyr::left_join(productivity, by = "Sample_16S") %>%                          # Join 26 OTU absolute abundance counts with rest of metadata 
  dplyr::filter(Lake == "Muskegon" & Depth == "Surface") %>%                     # Filter out samples that have productivity data (Muskegon & Surface)
  dplyr::select(-c(Platform, samples, Lake)) %>%                                 # Remove unnecessary columns 
  mutate(Site = factor(Site, levels = c("MOT", "MDP", "MBR", "MIN"))) %>%        # Fix order of factor for lake station
  gather("OTU", "Abs_Abund", 2:pick_columns) %>%                                 # Gather only columns which represent OTU abs abundance counts, and put it in *long* format
  dplyr::left_join(tax, by = "OTU") %>%                                          # Add the taxonomic information for each OTU
  dplyr::filter(tot_bacprod < 90) %>%                                            # Remove outliers to match peter's analysis
  mutate(OTU_fraction_HNA = Abs_Abund/HNA.cells)                                 # Calculate the fraction that each individual OTU takes up within the HNA pool for each sample



# Plot the absolute abundance data for each OTU
ggplot(AbsAbund_hna_otus, aes(x = reorder(OTU, Phylum), y = log10(Abs_Abund), fill = Phylum, color = Phylum)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 3) +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  ylab("log10(Abundance)") +
  #facet_grid(.~Phylum, scale = "free") +
  ggtitle("HNA OTUs from HNA Prediction") + 
  guides(fill = guide_legend(ncol=3),
         color = guide_legend(ncol=3)) +
  theme(legend.position = "bottom", #c(0.33, 0.85),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))

# Plot the fraction of the HNA pool for each OTU
ggplot(AbsAbund_hna_otus, aes(x = reorder(OTU, Phylum), y = OTU_fraction_HNA, fill = Phylum, color = Phylum)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 3) +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  ylab("Fraction of HNA Cells") +
  ggtitle("HNA OTUs from HNA Prediction") + 
  guides(fill = guide_legend(ncol=2),
         color = guide_legend(ncol=2)) +
  theme(legend.position = c(0.33, 0.85),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))
```

# Proportion of HNA Pool?

### What proportion of the HNA pool is taken up by the 26 OTUs?

```{r propostion-sum-abundance, fig.height=4, fig.width=3, fig.align='center'}
# Calculate the sum of the HNA pool, the max of the HNA pool, and the median/mean
frac_HNA_stats_AbsAbund_hna_otus <- AbsAbund_hna_otus %>%                        # Take the dataframe from above
  dplyr::select(Sample_16S, OTU, OTU_fraction_HNA, Abs_Abund, HNA.cells) %>%   # Select only relevant columns
  group_by(Sample_16S) %>%                                                     # Make a data frame for each of the individual samples
  summarise(sum_fracHNA = sum(OTU_fraction_HNA),                               # Calculate the sum of the fraction of each OTU from the HNA pool (total HNA pool represented by each OTU)
            max_fracHNA = max(OTU_fraction_HNA),                               # Calculate the max the the line above
            median_fracHNA = median(OTU_fraction_HNA),                         # Calculate the median of 2 lines above
            mean_fracHNA = mean(OTU_fraction_HNA),                             # Calculate the mean of 3 lines above
            sum_abs_abund = sum(Abs_Abund)) %>%                                # What's the sum of the absolute abundance of the 26 OTUs within each sample? 
  mutate(All_Samples = "AllSamps_26_OTUs")


# Plot the variation in the sum of the HNA fraction with points/boxplot
ggplot(frac_HNA_stats_AbsAbund_hna_otus, 
       aes(y = sum_fracHNA, x = All_Samples, color = "All_Samples", fill = "All_Samples")) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) + 
  geom_point(size = 3, position = position_jitterdodge()) +
  ggtitle("HNA OTUs from HNA Prediction") + 
  xlab("All Samples") + scale_color_manual(values = "black") +
  scale_fill_manual(values = "black") +
  geom_abline(intercept = 1, slope = 0, color = "grey") +                                   # Draw a line at 1 
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, max(frac_HNA_stats_AbsAbund_hna_otus$sum_fracHNA) + 0.2), 
                     breaks = seq(0,  max(frac_HNA_stats_AbsAbund_hna_otus$sum_fracHNA) + 0.2, by = 0.1)) +
  ylab("\n Sum(Abundance/HNA.cells)") + xlab("Sample") +
  theme(legend.position = "none", axis.text.x = element_blank())
```


# Sum OTUs vs HNA
```{r pred-HNA, fig.height=4, fig.width=4, fig.align='center'}
all_data <- inner_join(frac_HNA_stats_AbsAbund_hna_otus, productivity, by = "Sample_16S") %>%    # Combine calculated stats above with rest of metadata
  dplyr::select(-c(Platform, All_Samples, samples, Lake, Fraction, Depth)) %>%                  # Remove futile columns
  mutate(pred_totHNA_counts = sum_fracHNA*HNA.cells)                                            # Sanity Check:  Back calculate the number of HNA cells 

# Plot the correlation between the HNA counts from the flow cytometer and the sum of the counts from the 26 OTUs
ggplot(all_data, aes(x = sum_abs_abund, y= HNA.cells)) +
  geom_point(size = 3) + ylab("HNA Cell Count") + 
  xlab("Sum(Abs_Abund of 26 OTUs)") + 
  geom_abline(intercept = 0, slope = 1) +                                   # Draw a 1:1 line 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))
```



```{r otu-prod, fig.height=4, fig.width=10, eval = FALSE, echo = FALSE}

# Linear Regressions with the 3 OTUs pulled out in both analyses

# Simple OLS regression with Otu000025
lm_Otu000025 <- lm(tot_bacprod ~ log10(Abs_Abund) , data = dplyr::filter(AbsAbund_OTUs_26, OTU == "Otu000025"))

plot_OTU25 <- ggplot(dplyr::filter(AbsAbund_OTUs_26, OTU == "Otu000025"), 
       aes(y = tot_bacprod,x = log10(Abs_Abund), fill = Phylum, color = Phylum)) +
  geom_point(size = 3) + ggtitle("Otu000025") +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  xlab("log10(Abundance)") +
  geom_smooth(method = "lm", color = "#FFC543") +
  ylab("Total Production (ug C/L/hr)") + 
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  annotate("text", x = 4.45, y=60, color = "#FFC543", fontface = "bold",
           label = paste("R2 =", round(summary(lm_Otu000025)$adj.r.squared, digits = 2), "\n", 
                         "p =", round(unname(summary(lm_Otu000025)$coefficients[,4][2]), digits = 3)))  


# Simple OLS regression with Otu000041
lm_Otu000041 <- lm(tot_bacprod ~ log10(Abs_Abund), 
                   data = dplyr::filter(AbsAbund_OTUs_26, OTU == "Otu000041" & Abs_Abund > 0))  # Remove sample that has count of 0

plot_OTU41 <- ggplot(dplyr::filter(AbsAbund_OTUs_26, OTU == "Otu000041"), 
       aes(y = tot_bacprod,x = log10(Abs_Abund), fill = Phylum, color = Phylum)) +
  geom_point(size = 3) +  ggtitle("Otu000041") +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  xlab("log10(Abundance)") +
  geom_smooth(method = "lm", color = "#562258") +
  ylab("Total Production (ug C/L/hr)") + 
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  annotate("text", x = 5.1, y=60, color = "#562258", fontface = "bold",
         label = paste("R2 =", round(summary(lm_Otu000041)$adj.r.squared, digits = 2), "\n", 
                       "p =", round(unname(summary(lm_Otu000041)$coefficients[,4][2]), digits = 3)))  


# Simple OLS regression with Otu000176
lm_Otu000176 <- lm(tot_bacprod ~ log10(Abs_Abund), 
                   data = dplyr::filter(AbsAbund_OTUs_26, OTU == "Otu000176" & Abs_Abund > 0))

plot_OTU176 <- ggplot(dplyr::filter(AbsAbund_OTUs_26, OTU == "Otu000176"), 
       aes(y = tot_bacprod,x = log10(Abs_Abund), fill = Phylum, color = Phylum)) +
  geom_point(size = 3) + ggtitle("Otu000176") +
  scale_color_manual(values = phylum_colors) +
  scale_fill_manual(values = phylum_colors) +
  xlab("log10(Abundance)") +
  geom_smooth(method = "lm", color = "#FF2151") +
  ylab("Total Production (ug C/L/hr)") + 
  theme(legend.position = "bottom",
        legend.title = element_blank())+
  annotate("text", x = 3.5, y=60, color = "#FF2151", fontface = "bold",
         label = paste("R2 =", round(summary(lm_Otu000176)$adj.r.squared, digits = 2), "\n", 
                       "p =", round(unname(summary(lm_Otu000176)$coefficients[,4][2]), digits = 3)))  

# Plot all 3 plots at once
plot_grid(plot_OTU25, plot_OTU176, plot_OTU41, nrow = 1, ncol = 3,
          labels = c("A", "B", "C"))
```

**Note that All samples with an absolute abundance of 0 were left out of the above OLS linear model calculations.**

