---
author: "Marian L. Schmidt"
date: "April 28th, 2017"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 2
---



```{r setup, include=FALSE}
# For width of code chunks and scroll bar 
options(width=250)

knitr::opts_chunk$set(eval = TRUE, 
                      echo = TRUE,
                      include = TRUE,
                      warning = FALSE,
                      collapse = FALSE,
                      message = FALSE,
                      error = TRUE,
                      fig.align = "center")
```



### Load the necessary libraries and set colors 
```{r load-libraries}
################################# LOAD LIBRARIES ############################################
library(tidyverse)
library(cowplot)
source("../set_colors.R")
source("../Functions.R")
```


### Load in the data
```{r load-data}
#################################### LOAD DATA ##############################################
# Read in the absolute abundance data 
absolute_otu <- read.table(file="../../../data/Chloroplasts_removed/nochloro_absolute_otu.tsv", header = TRUE) # Absolute OTU abundance table 

# Read in the taxonomy data 
tax <- read.table(file="../../../data/Chloroplasts_removed/nochloro_taxonomy_otu.tsv", header = TRUE) %>%
  tibble::rownames_to_column() %>%
  dplyr::rename(Kingdom = Rank1, Phylum = Rank2, Class = Rank3, Order = Rank4, 
                Family = Rank5, Genus = Rank6, Species = Rank7, OTU = rowname) # Fix the Taxonomy

# Replace the phylum Proteobacteria with the class level
Phylum <- as.character(tax$Phylum)
Class <- as.character(tax$Class)

for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Proteobacteria"){
    Phylum[i] <- Class[i]
  } 
}

# Overwrite the Phylum level with the new phylum classification
tax$Phylum <- Phylum # Add the new phylum level data back to phy

# Read in the productivity and flow cytometry data 
prod_fcm_data <- read.table(file = "../../../data/Chloroplasts_removed/productivity_data.tsv", header = TRUE) # Metadata file


# Read in the metadata for the 
inland_fcm_data <- read.csv2("../../../data/inland.sam.csv", stringsAsFactors = FALSE) %>%
  mutate(Sample_16S = paste(substr(SampleID, 1,3), substr(SampleID, 5,7), substr(SampleID, 9,9), sep = "")) %>%
  dplyr::filter(Fraction == "Free")
```


# Load in selected OTUs

## How many OTUs are present in all each of the analyses?

```{r load-OTUs}
# Load in the 4 files of OTUs

# Abundance filtering based on All of the sample  data 
## Abundance cutoff = 1%
otus_stand_0.01 <- read.csv("HNA_selectedOTUs_stand_abun0.01_R20.959411030714.csv", header = FALSE) %>% rename(OTU = V1, score = V2)
## How many OTUs are in this dataset?
nrow(otus_stand_0.01)

## Abundance cutoff = 0.1%
otus_stand_0.001 <- read.csv("HNA_selectedOTUs_stand_abun0.001_R20.969251756656.csv", header = FALSE) %>% rename(OTU = V1, score = V2)
## How many OTUs are in this dataset?
nrow(otus_stand_0.001)

# Abundance filtering based on the Productivity data
## Abundance cutoff = 1%
otus_prod_0.001 <- read.csv("HNA_selectedOTUs_prod_stand_abun0.001_R20.928038799058.csv", header = FALSE) %>% rename(OTU = V1, score = V2) 
## How many OTUs are in this dataset?
nrow(otus_prod_0.001)

## Abundance cutoff = 0.1%
otus_prod_0.01 <- read.csv("HNA_selectedOTUs_prod_stand_abun0.01_R20.87968307489.csv", header = FALSE) %>% rename(OTU = V1, score = V2)
## How many OTUs are in this dataset?
nrow(otus_prod_0.01)
```


## How many are present in the different types of analyses?

```{r otu-intersection}
## Standardized analysis with all samples?
length(intersect(otus_stand_0.001$OTU, otus_stand_0.01$OTU))
## Standardized analysis with only productivity samples?
length(intersect(otus_prod_0.001$OTU, otus_prod_0.01$OTU))

# How many OTUs are shared between all 4 approaches?
otus_in_all_4 <- Reduce(intersect, list(otus_stand_0.001$OTU, otus_stand_0.01$OTU, otus_prod_0.001$OTU, otus_prod_0.01$OTU))
length(otus_in_all_4)

# What is the taxonomy of the OTUs that are shared between all 4 approaches?
tax %>% dplyr::filter(OTU %in% otus_in_all_4)
```



# Analysis of HNA OTUs that are found within **BOTH** of the models

```{r intersect-HNA, fig.width=14, fig.height=12}
## Standardized analysis with all samples?
stand_OTUs <- intersect(otus_stand_0.001$OTU, otus_stand_0.01$OTU)

# Put all the data together into one dataframe with only the important OTUs
AbsAbund_otus_stand_intersect <- combine_OTU_data(absolute_otu_table = absolute_otu, otu_vector_names = stand_OTUs, 
                                             productivity_fcm_data = prod_fcm_data, taxonomy_table = tax)

length(unique(AbsAbund_otus_stand_intersect$OTU))


ggplot(AbsAbund_otus_stand_intersect, aes(x = Abs_Abund, y = HNA.cells)) + 
  geom_point() + facet_wrap(~ OTU, scale = "free") 
```




# 1% Cutoff from ALL samples 

```{r all-1%-cutoff, dependson = "load-data", fig.height=4, fig.width=3, fig.align='center'}
# What is the taxonomy of the OTUs?
tax %>% dplyr::filter(OTU %in% otus_stand_0.01$OTU)

# Put all the data together into one dataframe with only the important OTUs
AbsAbund_otus_stand_0.01 <- combine_OTU_data(absolute_otu_table = absolute_otu, otu_vector_names = as.vector(otus_stand_0.01$OTU), 
                                             productivity_fcm_data = prod_fcm_data, taxonomy_table = tax)


HNA_frac_otus_stand_0.01 <- calc_fraction_HNA(AbsAbund_OTUs = AbsAbund_otus_stand_0.01) %>%
  arrange(sum_fracHNA) %>%
  left_join(inland_fcm_data, by = "Sample_16S") %>%
  dplyr::select(Sample_16S, sum_fracHNA, sum_abs_abund, All_Samples, Station, Season, Lake, Depth)


# Plot the variation in the sum of the HNA fraction with points/boxplot
plot1 <-  ggplot(HNA_frac_otus_stand_0.01, 
       aes(y = sum_fracHNA, x = All_Samples, color = "All_Samples", fill = "All_Samples")) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +   geom_point(size = 3, position = position_jitterdodge()) +
  ggtitle("All\n 1% Cutoff") + ylab("\n Sum(Abundance/HNA.cells)") + xlab("Sample") +
  scale_color_manual(values = "black") + scale_fill_manual(values = "black") +
  geom_abline(intercept = 1, slope = 0, color = "red") +                                   # Draw a line at 1 
  scale_y_continuous(expand = c(0,0),limits = c(0, 2.5), breaks = seq(0, 2.5, by = 0.5)) +
  theme(legend.position = "none", axis.text.x = element_blank())

# To see the samples at the bottom of the plot?
head(HNA_frac_otus_stand_0.01)
```

# 0.1% Cutoff from ALL samples 

```{r all-0.1%-cutoff, dependson = "load-data", fig.height=4, fig.width=3, fig.align='center'}
# What is the taxonomy of the OTUs?
tax %>% dplyr::filter(OTU %in% otus_stand_0.001$OTU)

# Put all the data together into one dataframe with only the important OTUs
AbsAbund_otus_stand_0.001 <- combine_OTU_data(absolute_otu_table = absolute_otu, otu_vector_names = as.vector(otus_stand_0.001$OTU), 
                                             productivity_fcm_data = prod_fcm_data, taxonomy_table = tax)

# Calculate the Sum of the HNA fractions
HNA_frac_otus_stand_0.001 <- calc_fraction_HNA(AbsAbund_OTUs = AbsAbund_otus_stand_0.001) %>%
  arrange(sum_fracHNA) %>%
  left_join(inland_fcm_data, by = "Sample_16S") %>%
  dplyr::select(Sample_16S, sum_fracHNA, sum_abs_abund, All_Samples, Station, Season, Lake, Depth)

# Plot the variation in the sum of the HNA fraction with points/boxplot
plot2 <- ggplot(HNA_frac_otus_stand_0.001, 
       aes(y = sum_fracHNA, x = All_Samples, color = "All_Samples", fill = "All_Samples")) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +   geom_point(size = 3, position = position_jitterdodge()) +
  ggtitle("All\n 0.1% Cutoff") + ylab("\n Sum(Abundance/HNA.cells)") + xlab("Sample") +
  scale_color_manual(values = "black") + scale_fill_manual(values = "black") +
  geom_abline(intercept = 1, slope = 0, color = "red") +                                   # Draw a line at 1 
  scale_y_continuous(expand = c(0,0),limits = c(0, 2.5), breaks = seq(0, 2.5, by = 0.5)) +
  theme(legend.position = "none", axis.text.x = element_blank())

# To see the samples at the bottom of the plot?
head(HNA_frac_otus_stand_0.001)
```



# Is the HNA percent higher/lower in the hypolimnion of productvite stratified inland lakes?


```{r HNA_percent-hypolimnia, fig.width=8, fig.height=4, fig.align='center'}
dat <- prod_fcm_data %>%
  dplyr::filter(Lake == "Inland" & Season == "Summer") %>%
  mutate(HNA_percent = HNA.cells/Total.cells)

p1 <- ggplot(dat, aes(y = HNA_percent, x = Depth, color = Depth, shape = Season, fill = Depth)) + 
  geom_jitter(size = 3) + 
  ggtitle("All Summer Samples") +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  theme(axis.title.x = element_blank(), 
        legend.position = "none")
  
erp <- dplyr::filter(dat, Depth == "Deep") %>% 
  dplyr::select(Site,HNA_percent, Depth) %>%
  arrange(HNA_percent) 

# Fix the order of the x axis
erp$Site <- factor(erp$Site, levels=unique(as.character(erp$Site)) )

# 
p2 <- ggplot(erp, aes(x = Site, y = HNA_percent, color = Depth)) + 
  geom_point(size = 3) +
  ggtitle("Summer Hypolimnia") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        axis.title.x = element_blank(), legend.position = "none")

plot_grid(p1, p2, align = "h",
          nrow = 1, ncol = 2,
          labels = c("A", "B"),
          rel_widths = c(1, 2))
```

## The HNA percent is much higher in the hypolimnion of productvite stratified lakes!



# 1% Cutoff from Productivity samples only 

```{r prod-1%-cutoff, dependson = "load-data", fig.height=4, fig.width=3, fig.align='center'}
# What is the taxonomy of the OTUs?
tax %>% dplyr::filter(OTU %in% otus_prod_0.01$OTU)

# Put all the data together into one dataframe with only the important OTUs
AbsAbund_otus_prod_0.01 <- combine_OTU_data(absolute_otu_table = absolute_otu, otu_vector_names = as.vector(otus_prod_0.01$OTU), 
                                             productivity_fcm_data = prod_fcm_data, taxonomy_table = tax) %>%
        dplyr::filter(Lake == "Muskegon" & Depth == "Surface")               # Filter out samples that have productivity data (Muskegon & Surface)


# Calculate the Sum of the HNA fractions
HNA_frac_otus_prod_0.01 <- calc_fraction_HNA(AbsAbund_OTUs = AbsAbund_otus_prod_0.01)

# Plot the variation in the sum of the HNA fraction with points/boxplot
plot3 <- ggplot(HNA_frac_otus_prod_0.01, 
       aes(y = sum_fracHNA, x = All_Samples, color = "All_Samples", fill = "All_Samples")) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +   geom_point(size = 3, position = position_jitterdodge()) +
  ggtitle("Prod\n 1% Cutoff") + ylab("\n Sum(Abundance/HNA.cells)") + xlab("Sample") +
  scale_color_manual(values = "black") + scale_fill_manual(values = "black") +
  geom_abline(intercept = 1, slope = 0, color = "red") +                                   # Draw a line at 1 
  scale_y_continuous(expand = c(0,0),limits = c(0, 2.5), breaks = seq(0, 2.5, by = 0.5)) +
  theme(legend.position = "none", axis.text.x = element_blank())
```


# 0.1% Cutoff from Productivity samples only 

```{r prod-0.1%-cutoff, dependson = "load-data", fig.height=4, fig.width=3, fig.align='center'}
# What is the taxonomy of the OTUs?
tax %>% dplyr::filter(OTU %in% otus_prod_0.001$OTU)

# Put all the data together into one dataframe with only the important OTUs
AbsAbund_otus_prod_0.001 <- combine_OTU_data(absolute_otu_table = absolute_otu, otu_vector_names = as.vector(otus_prod_0.001$OTU), 
                                             productivity_fcm_data = prod_fcm_data, taxonomy_table = tax) %>%
      dplyr::filter(Lake == "Muskegon" & Depth == "Surface")               # Filter out samples that have productivity data (Muskegon & Surface)


# Calculate the Sum of the HNA fractions
HNA_frac_otus_prod_0.001 <- calc_fraction_HNA(AbsAbund_OTUs = AbsAbund_otus_prod_0.001)

# Plot the variation in the sum of the HNA fraction with points/boxplot
plot4 <- ggplot(HNA_frac_otus_prod_0.001, 
       aes(y = sum_fracHNA, x = All_Samples, color = "All_Samples", fill = "All_Samples")) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +   geom_point(size = 3, position = position_jitterdodge()) +
  ggtitle("Prod\n 0.1% Cutoff") + ylab("\n Sum(Abundance/HNA.cells)") + xlab("Sample") +
  scale_color_manual(values = "black") + scale_fill_manual(values = "black") +
  geom_abline(intercept = 1, slope = 0, color = "red") +                                   # Draw a line at 1 
  scale_y_continuous(expand = c(0,0),limits = c(0, 2.5), breaks = seq(0, 2.5, by = 0.5)) +
  theme(legend.position = "none", axis.text.x = element_blank())
```

# All Plots 
```{r all-plots, dependson = c("all-1%-cutoff", "all-0.1%-cutoff", "prod-1%-cutoff", "prod-0.1%-cutoff"), fig.height=4, fig.width=10, fig.align='center'}
plot_grid(plot1, plot2, plot3, plot4, align = "h",
          labels = c("A", "B", "C", "D"),
          nrow = 1, ncol = 4)

```



